### 2023-02-21 219-2023-analysis/manuscript.qmd
use osf storage, update preprocess, add quarto doc

---
title: "Replication of Brydevall et al. 2018"
author: "Josh de Leeuw"
format: html
editor: visual
---

```{r}
#| label: Load libraries
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(readr)
```

## Results

### Behavioral

```{r}
#| label: Load the Behavioral Data
#| echo: false
#| message: false

data.behavioral <- read_csv("data/preprocessed/behavioral.csv")

```

```{r}
#| label: Attention check calculations
#| echo: false

attention.check.responses <- data.behavioral %>%
  filter(task=="catch", phase=="test", card == catch_n) %>%
  group_by(subject_id) %>%
  summarize(successes = sum(rt < 1500), mean_rt = mean(rt))

attention.check.overall.m <- attention.check.responses %>% pull(successes) %>% mean()
attention.check.overall.sd <- attention.check.responses %>% pull(successes) %>%
  sd()
attention.check.rt.m <- attention.check.responses %>% pull(mean_rt) %>%
  mean()
attention.check.rt.sd <- attention.check.responses %>% pull(mean_rt) %>%
  sd()

attention.check.failures <- attention.check.responses %>% filter(successes < 3)
```

There were five attention check trials, one during each block, throughout the experiment. The mean number of correct responses was `r attention.check.overall.m` (*SD* = `r attention.check.overall.sd`) and the average response time was We excluded `r nrow(attention.check.failures)` participants for failing more than 2 attention checks.

### EEG

```{r}
#| label: Load EEG data
#| echo: false
#| message: false

data.eeg <- read_csv('data/preprocessed/eeg.csv')

```

```{r}
#| label: Filter EEG data
#| echo: false

data.eeg.filtered <- data.eeg %>%
  filter(good_segment == TRUE) %>%
  filter(electrode %in% c("Cz", "Fz"))

```

```{r}
#| label: Compute RPEs and IPEs
#| echo: false
#| message: false

hand.information <- data.behavioral %>% 
  filter(phase=='test', task=='reveal') %>%
  mutate(subject = subject_id) %>%
  select(-subject_id) %>%
  group_by(subject) %>%
  mutate(hand_id = rep(1:80, each=5)) %>%
  mutate(losses_so_far = (card-1) - wins_so_far) %>%
  select(subject, hand_id, card, card_value, wins_so_far, losses_so_far, outcome)

hand.information.with.rpe <- hand.information %>%
  mutate(
    wp_before = pbinom(2-losses_so_far, 5-(card-1), 0.5), 
    wp_after = pbinom(2-(losses_so_far+abs(1-card_value)), 5-card, 0.5), 
    rpe = wp_after-wp_before)

entropy <- function(p){
  return(ifelse(p == 0 | p == 1, 0, -p*log2(p) - (1-p)*log2(1-p)))
}

hand.information.with.rpe.and.ipe <- hand.information.with.rpe %>%
  mutate(wp_alternative = pbinom(2-(losses_so_far+card_value), 5-card, 0.5)) %>%
  mutate(information = entropy(wp_before) - entropy(wp_after),
         information_expected = ((entropy(wp_before) - entropy(wp_after)) + (entropy(wp_before) - entropy(wp_alternative))) / 2,
         ipe = information - information_expected)

data.eeg.filtered.rpe.ipe <- data.eeg.filtered %>%
  mutate(card = card_id) %>%
  select(-card_id) %>%
  left_join(hand.information.with.rpe.and.ipe, by=c("subject", "hand_id", "card"))

```