### 2023-02-23 219-2023-analysis/manuscript.qmd
add analysis work

---
title: "Replication of Brydevall et al. 2018"
author: "Josh de Leeuw"
format: 
  html:
    df-print: kable
editor: visual
---

```{r}
#| label: Load libraries
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(readr)
library(ez)
```

## Results

### Behavioral

```{r}
#| label: Load the Behavioral Data
#| echo: false
#| message: false

data.behavioral <- read_csv("data/preprocessed/behavioral.csv")

```

```{r}
#| label: Attention check calculations
#| echo: false

attention.check.responses <- data.behavioral %>%
  filter(task=="catch", phase=="test", card == catch_n) %>%
  group_by(subject_id) %>%
  summarize(successes = sum(rt < 1500), mean_rt = mean(rt))

attention.check.overall.m <- attention.check.responses %>% pull(successes) %>% mean()
attention.check.overall.sd <- attention.check.responses %>% pull(successes) %>%
  sd()
attention.check.rt.m <- attention.check.responses %>% pull(mean_rt) %>%
  mean() %>% round(0)
attention.check.rt.sd <- attention.check.responses %>% pull(mean_rt) %>%
  sd() %>% round(0)

attention.check.failures <- attention.check.responses %>% filter(successes < 3)
```

There were five attention check trials, one during each block, throughout the experiment. The mean number of correct responses was `r attention.check.overall.m` (*SD* = `r attention.check.overall.sd`) and the average response time was `r attention.check.rt.m`ms (*SD* = `r attention.check.rt.sd`). As per our pre-registered criteria, we excluded `r nrow(attention.check.failures)` participants for failing more than 2 attention checks (defined as a response time greater than 1,500ms).

### EEG

```{r}
#| label: Load EEG data
#| echo: false
#| message: false

data.eeg <- read_csv('data/preprocessed/eeg.csv')

```

```{r}
#| label: Filter EEG data
#| echo: false

data.eeg.filtered <- data.eeg %>%
  filter(good_segment == TRUE) %>%
  filter(electrode %in% c("Cz", "Fz"))

```

```{r}
#| label: Compute RPEs and IPEs
#| echo: false
#| message: false

hand.information <- data.behavioral %>% 
  filter(phase=='test', task=='reveal') %>%
  mutate(subject = subject_id) %>%
  select(-subject_id) %>%
  group_by(subject) %>%
  mutate(hand_id = rep(1:80, each=5)) %>%
  mutate(losses_so_far = (card-1) - wins_so_far) %>%
  select(subject, hand_id, card, card_value, wins_so_far, losses_so_far, outcome)

hand.information.with.rpe <- hand.information %>%
  mutate(
    wp_before = pbinom(2-losses_so_far, 5-(card-1), 0.5), 
    wp_after = pbinom(2-(losses_so_far+abs(1-card_value)), 5-card, 0.5), 
    rpe = wp_after-wp_before)

entropy <- function(p){
  return(ifelse(p == 0 | p == 1, 0, -p*log2(p) - (1-p)*log2(1-p)))
}

hand.information.with.rpe.and.ipe <- hand.information.with.rpe %>%
  mutate(wp_alternative = pbinom(2-(losses_so_far+card_value), 5-card, 0.5)) %>%
  mutate(information = entropy(wp_before) - entropy(wp_after),
         information_expected = ((entropy(wp_before) - entropy(wp_after)) + (entropy(wp_before) - entropy(wp_alternative))) / 2,
         ipe = information - information_expected) %>%
  mutate(rpe_type = case_when(
    rpe > 0 ~ "Positive",
    rpe < 0 ~ "Negative",
    rpe == 0 ~ "None")) %>%
  mutate(ipe_type = case_when(
    ipe > 0 ~ "Positive",
    ipe < 0 ~ "Negative",
    ipe == 0 ~ "None"))

data.eeg.filtered.rpe.ipe <- data.eeg.filtered %>%
  mutate(card = card_id) %>%
  select(-card_id) %>%
  left_join(hand.information.with.rpe.and.ipe, by=c("subject", "hand_id", "card"))

```

```{r}
#| label: Define ERP time window
#| echo: false

frn.window <- 200:350
```

```{r}
#| label: Count usable segments for RPEs
#| echo: false
#| message: false

rpe.segments <- data.eeg.filtered.rpe.ipe %>%
  filter(rpe_type != "None") %>%
  group_by(subject, electrode, hand_id, card, rpe_type) %>%
  summarize(n = n()) %>%
  group_by(subject, electrode, rpe_type) %>%
  summarize(n = n())
```

```{r}
#| label: Count usable segments for IPEs
#| echo: false
#| message: false

ipe.segments <- data.eeg.filtered.rpe.ipe %>%
  filter(ipe_type != "None") %>%
  group_by(subject, electrode, hand_id, card, ipe_type) %>%
  summarize(n = n()) %>%
  group_by(subject, electrode, ipe_type) %>%
  summarize(n = n())
```

We pre-registered an inclusion criteria that a participant must have at least 20 usable epochs of data of a particular event type in order to be included in the corresponding analyses. The minimum number of usable epochs for the RPE analysis in an individual subject was `r rpe.segments %>% pull(n) %>% min()`, *M*=`r rpe.segments %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r rpe.segments %>% pull(n) %>% sd() %>% round(0)`. The minimum number of usable epochs for the IPE analysis in an individual subject was `r ipe.segments %>% pull(n) %>% min()`, *M*=`r ipe.segments %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% pull(n) %>% sd() %>% round(0)`. Therefore, no participants needed to be excluded from the following analyses.

```{r}
#| label: Generate data for Grand Averages
#| echo: false
#| message: false

rpe.erp.data.subject <- data.eeg.filtered.rpe.ipe %>%
  filter(rpe_type != "None") %>%
  group_by(subject, electrode, t, rpe_type) %>%
  summarize(m.subject = mean(v))

rpe.erp.data.group <- rpe.erp.data.subject %>%
  group_by(electrode, t, rpe_type) %>%
  summarize(m = mean(m.subject), se = sd(m.subject)/sqrt(n())) %>%
  mutate(error_type = "Reward Prediction Error", error_sign = rpe_type)

ipe.erp.data.subject <- data.eeg.filtered.rpe.ipe %>%
  filter(ipe_type != "None") %>%
  group_by(subject, electrode, t, ipe_type) %>%
  summarize(m.subject = mean(v))

ipe.erp.data.group <- ipe.erp.data.subject %>%
  group_by(electrode, t, ipe_type) %>%
  summarize(m = mean(m.subject), se = sd(m.subject)/sqrt(n())) %>%
  mutate(error_type = "Information Prediction Error", error_sign = ipe_type)

erp.data <- rpe.erp.data.group %>% bind_rows(ipe.erp.data.group) %>%
  select(-rpe_type, -ipe_type)

```

```{r}
#| label: fig-grand-averages
#| echo: false
#| fig-cap: Grand average ERPs for both electrode sites (Cz, Fz) and both error types (Information Prediction and Reward Prediction). The critical time window for our analysis, 200-350ms, is highlighted.

ggplot(erp.data, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=error_sign, fill=error_sign))+
  facet_grid(error_type ~ electrode)+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  annotate("rect", xmin=min(frn.window), xmax=max(frn.window), ymin=-Inf, ymax=Inf, alpha=0.4)+
  geom_ribbon(color=NA, alpha = 0.2)+
  geom_line()+
  scale_y_reverse()+
  scale_color_brewer(type="qual", palette = "Set1", name="Error Direction")+
  scale_fill_brewer(type="qual", palette = "Set1", name="Error Direction")+
  labs(x="Time since card value revealed (ms)", y="Microvolts")+
  theme_minimal()+
  theme(panel.grid = element_blank())

```

```{r}
#| label: Compute ANOVA data
#| echo: false
#| message: false

rpe.anova.data <- rpe.erp.data.subject %>%
  filter(t %in% frn.window) %>%
  group_by(subject, electrode, rpe_type) %>%
  summarize(M = mean(m.subject))

ipe.anova.data <- ipe.erp.data.subject %>%
  filter(t %in% frn.window) %>%
  group_by(subject, electrode, ipe_type) %>%
  summarize(M = mean(m.subject))
```

```{r}
#| label: RPE ANOVA
#| echo: false
#| warning: false

rpe.anova <- ezANOVA(rpe.anova.data, dv=M, wid=subject, within=c(electrode, rpe_type))

rpe.anova$ANOVA
```

```{r}
#| label: IPE ANOVA
#| echo: false
#| warning: false

ipe.anova <- ezANOVA(ipe.anova.data, dv=M, wid=subject, within=c(electrode, ipe_type))

ipe.anova$ANOVA
```