### 2023-04-04 219-2023-analysis/manuscript.qmd
analysis with all 40 subjects!

---
title: "Replication of Brydevall et al. 2018"
author: "Josh de Leeuw"
format: 
  html:
    df-print: kable
editor: visual
---

```{r}
#| label: Load libraries
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(readr)
library(ez)
library(tidyr)
library(osfr)
```

```{r}
#| label: Get preprocessed data from OSF
#| include: false
#| echo: false

osf_retrieve_node("2pxf4") %>%
  osf_ls_files() %>%
  osf_download(path="data/preprocessed/", conflicts="skip")
```

## Methods

```{r}
#| label: Load the Behavioral Data
#| echo: false
#| message: false

data.behavioral <- read_csv("data/preprocessed/behavioral.csv")
```

### Participants

```{r}
#| label: Count participants
#| echo: false

n.participants <- data.behavioral %>% pull(subject) %>% unique() %>% length()
```

`r n.participants` Vassar College students participated in this study. All participants were right-handed.

## Results

### Behavioral

```{r}
#| label: Attention check calculations
#| echo: false

attention.check.responses <- data.behavioral %>%
  filter(task=="catch", phase=="test", card_id == catch_n) %>%
  group_by(subject) %>%
  summarize(successes = sum(rt < 1500, na.rm=TRUE), mean_rt = mean(rt, na.rm=TRUE))

attention.check.tally <- attention.check.responses %>% group_by(successes) %>% summarize(n=n())

attention.check.rt.m <- attention.check.responses %>% pull(mean_rt) %>%
  mean() %>% round(0)
attention.check.rt.sd <- attention.check.responses %>% pull(mean_rt) %>%
  sd() %>% round(0)

attention.check.failures <- attention.check.responses %>% filter(successes < 3)
```

There were five attention check trials, one during each block, throughout the experiment. `r attention.check.tally %>% filter(successes==5) %>% pull(n)` of `r n.participants` responded in time to all 5 attention checks, and `r attention.check.tally %>% filter(successes==4) %>% pull(n)` responded correctly to 4 of the 5 attention checks. The average response time was `r attention.check.rt.m`ms (*SD* = `r attention.check.rt.sd`). As per our pre-registered criteria, we excluded `r nrow(attention.check.failures)` participants for failing more than 2 attention checks (defined as a response time greater than 1,500ms).

```{r}
#| label: Compute RPEs and IPEs
#| echo: false
#| message: false

hand.information <- data.behavioral %>% 
  filter(phase=='test', task=='reveal') %>%
  group_by(subject) %>%
  mutate(hand_id = rep(1:80, each=5)) %>%
  ungroup() %>%
  mutate(losses_so_far = (card_id-1) - wins_so_far) %>%
  select(subject, hand_id, card_id, card_value, wins_so_far, losses_so_far, outcome)

hand.information.with.rpe <- hand.information %>%
  mutate(
    wp_before = pbinom(2-losses_so_far, 5-(card_id-1), 0.5), 
    wp_after = pbinom(2-(losses_so_far+(1-card_value)), 5-card_id, 0.5), 
    rpe = wp_after-wp_before)

entropy <- function(p){
  return(ifelse(p == 0 | p == 1, 0, -p*log2(p) - (1-p)*log2(1-p)))
}

hand.information.with.rpe.and.ipe <- hand.information.with.rpe %>%
  mutate(wp_alternative = pbinom(2-(losses_so_far+card_value), 5-card_id, 0.5)) %>%
  mutate(information = entropy(wp_before) - entropy(wp_after),
         information_alt = entropy(wp_before) - entropy(wp_alternative),
         information_expected = (information + information_alt) / 2,
         ipe = information - information_expected) %>%
  mutate(rpe_type = case_when(
    rpe > 0 ~ "Positive",
    rpe < 0 ~ "Negative",
    rpe == 0 ~ "None")) %>%
  mutate(ipe_type = case_when(
    ipe > 0 ~ "Positive",
    ipe < 0 ~ "Negative",
    ipe == 0 ~ "None"))

hand.information.minimal.set <- hand.information.with.rpe.and.ipe %>%
  select(subject, hand_id, card_id, rpe, ipe, rpe_type, ipe_type)
```

### EEG

```{r}
#| label: Load EEG data
#| echo: false
#| message: false

data.eeg.filtered <- read_csv('data/preprocessed/eeg.csv') %>%
  filter(good_segment == TRUE, electrode %in% c("Cz", "Fz"))

```

```{r}
#| label: Merge RPE and IPE info into EEG data
#| echo: false
#| message: false

data.eeg.filtered.rpe.ipe <- data.eeg.filtered %>%
  left_join(hand.information.minimal.set, by=c("subject", "hand_id", "card_id"))

```

```{r}
#| label: Count usable segments for RPEs
#| echo: false
#| message: false

rpe.segments <- data.eeg.filtered.rpe.ipe %>%
  filter(rpe_type != "None") %>%
  group_by(subject, electrode, hand_id, card_id, rpe_type) %>%
  slice_head(n=1) %>%
  group_by(subject, electrode, rpe_type) %>%
  summarize(n=n())
```

```{r}
#| label: Count usable segments for IPEs
#| echo: false
#| message: false

ipe.segments <- data.eeg.filtered.rpe.ipe %>%
  filter(ipe_type != "None") %>%
  group_by(subject, electrode, hand_id, card_id, ipe_type) %>%
  slice_head(n=1) %>%
  group_by(subject, electrode, ipe_type) %>%
  summarize(n = n())
```

```{r}
#| label: Exclude bad EEG
#| echo: false

rpe.good.subjects <- rpe.segments %>%
  group_by(subject) %>%
  summarize(include = all(n >= 20)) %>%
  filter(include == TRUE) %>%
  pull(subject)

ipe.good.subjects <- ipe.segments %>%
  group_by(subject) %>%
  summarize(include = all(n >= 20)) %>%
  filter(include == TRUE) %>%
  pull(subject)
```

We pre-registered an inclusion criteria that a participant must have at least 20 usable epochs of data of a particular event type in order to be included in the corresponding analyses. `r rpe.good.subjects %>% length()` subjects were included in the RPE analysis and `r ipe.good.subjects %>% length()` subjects were included in the IPE analysis. The average number of usable epochs for the RPE analysis in an individual subject was `r rpe.segments %>% filter(subject %in% rpe.good.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r rpe.segments %>% filter(subject %in% rpe.good.subjects) %>% pull(n) %>% sd() %>% round(0)`. The average number of usable epochs for the IPE analysis was `r ipe.segments %>% filter(subject %in% ipe.good.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% filter(subject %in% ipe.good.subjects) %>% pull(n) %>% sd() %>% round(0)`.

```{r}
#| label: Define ERP time window
#| echo: false

frn.window <- 200:350
```

```{r}
#| label: Generate data for Grand Averages
#| echo: false
#| message: false

rpe.erp.data.subject <- data.eeg.filtered.rpe.ipe %>%
  filter(rpe_type != "None") %>%
  filter(subject %in% rpe.good.subjects) %>%
  group_by(subject, electrode, t, rpe_type) %>%
  summarize(m.subject = mean(v))

rpe.erp.data.group <- rpe.erp.data.subject %>%
  group_by(electrode, t, rpe_type) %>%
  summarize(m = mean(m.subject), se = sd(m.subject)/sqrt(n())) %>%
  mutate(error_type = "Reward Prediction Error", error_sign = rpe_type)

ipe.erp.data.subject <- data.eeg.filtered.rpe.ipe %>%
  filter(ipe_type != "None") %>%
  filter(subject %in% ipe.good.subjects) %>%
  group_by(subject, electrode, t, ipe_type) %>%
  summarize(m.subject = mean(v))

ipe.erp.data.group <- ipe.erp.data.subject %>%
  group_by(electrode, t, ipe_type) %>%
  summarize(m = mean(m.subject), se = sd(m.subject)/sqrt(n())) %>%
  mutate(error_type = "Information Prediction Error", error_sign = ipe_type)

erp.data <- rpe.erp.data.group %>% bind_rows(ipe.erp.data.group) %>%
  select(-rpe_type, -ipe_type)

```

```{r}
#| label: fig-grand-averages
#| echo: false
#| fig-cap: Grand average ERPs for both electrode sites (Cz, Fz) and both error types (Information Prediction and Reward Prediction). The critical time window for our analysis, 200-350ms, is highlighted.

ggplot(erp.data, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=error_sign, fill=error_sign))+
  facet_grid(error_type ~ electrode)+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  annotate("rect", xmin=min(frn.window), xmax=max(frn.window), ymin=-Inf, ymax=Inf, alpha=0.4)+
  geom_ribbon(color=NA, alpha = 0.2)+
  geom_line()+
  scale_y_reverse()+
  scale_color_brewer(type="qual", palette = "Set1", name="Error Direction")+
  scale_fill_brewer(type="qual", palette = "Set1", name="Error Direction")+
  labs(x="Time since card value revealed (ms)", y="Microvolts")+
  theme_minimal()+
  theme(panel.grid = element_blank())

```

```{r}
#| label: Difference wave summary data
#| echo: false
#| message: false

difference.wave.data.rpe <- rpe.erp.data.subject %>%
  pivot_wider(names_from = rpe_type, values_from = m.subject) %>%
  mutate(difference.subject = Positive - Negative)

difference.wave.summary.rpe <- difference.wave.data.rpe %>%
  group_by(electrode, t) %>%
  summarize(difference = mean(difference.subject), se=sd(difference.subject)/sqrt(n())) %>%
  mutate(error_type = "Reward Prediction Error")

difference.wave.data.ipe <- ipe.erp.data.subject %>%
  pivot_wider(names_from = ipe_type, values_from = m.subject) %>%
  mutate(difference.subject = Positive - Negative)

difference.wave.summary.ipe <- difference.wave.data.ipe %>%
  group_by(electrode, t) %>%
  summarize(difference = mean(difference.subject), se=sd(difference.subject)/sqrt(n())) %>%
  mutate(error_type = "Information Prediction Error")

difference.erp.data <- difference.wave.summary.rpe %>% bind_rows(difference.wave.summary.ipe)

```

```{r}
#| label: fig-difference-waves
#| echo: false
#| fig-cap: Grand average difference wave ERPs for both electrode sites (Cz, Fz). Values above 0 indicate that positive errors had a more positive ERP than negative errors. The critical time window for our analysis, 200-350ms, is highlighted.

ggplot(difference.erp.data, aes(x=t,y=difference, ymin=difference-se, ymax=difference+se, color=error_type, fill=error_type))+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  annotate("rect", xmin=min(frn.window), xmax=max(frn.window), ymin=-Inf, ymax=Inf, alpha=0.1)+
  geom_ribbon(color=NA, alpha=0.2)+
  geom_line()+
   scale_color_brewer(type="qual", palette = "Set2", name="Error Type")+
  scale_fill_brewer(type="qual", palette = "Set2", name="Error Type")+
  facet_grid(electrode~.)+
  labs(x="Time since card value revealed (ms)", y="Positive - Negative Prediction Errors\n(Microvolts)")+
  theme_minimal()+
  theme(panel.grid = element_blank())

```

Brydevall et al. reported that the ERPs in the 200-350ms window following positive prediction errors were more positive than the ERPs following negative prediction errors, for both information prediction errors and reward prediciton errors. Following our pre-registration, we ran two ANOVAs to see if there was a main effect of reward type (positive vs. negative). We included electrode (Cz vs. Fz) as a factor, as well as the interaction between reward type and electrode.

```{r}
#| label: Compute ANOVA data
#| echo: false
#| message: false

rpe.anova.data <- rpe.erp.data.subject %>%
  filter(t %in% frn.window) %>%
  group_by(subject, electrode, rpe_type) %>%
  summarize(M = mean(m.subject))

ipe.anova.data <- ipe.erp.data.subject %>%
  filter(t %in% frn.window) %>%
  group_by(subject, electrode, ipe_type) %>%
  summarize(M = mean(m.subject))
```

```{r}
#| label: RPE ANOVA
#| echo: false
#| warning: false

rpe.anova <- ezANOVA(rpe.anova.data, dv=M, wid=subject, within=c(electrode, rpe_type))

rpe.anova$ANOVA
```

The ANOVA table above shows that there is no significant effect of error type for reward prediction errors.

```{r}
#| label: IPE ANOVA
#| echo: false
#| warning: false

ipe.anova <- ezANOVA(ipe.anova.data, dv=M, wid=subject, within=c(electrode, ipe_type))

ipe.anova$ANOVA
```

However, there was a significant effect of error type for information prediction errors. This confirms the pattern that is evident in both @fig-grand-averages and @fig-difference-waves, where there is no clear visual difference between the ERPs for reward prediction errors, but there is a clear difference for information prediction errors.